---
title: "Opakapaka Simulation Report"
date: "September 17, 2025"
date-format: "MMMM D, YYYY"

# Authors and affiliations - try the authors structure instead
authors:
  - name: "Megumi Oshima, Marc Nadon, John Syslo,<br>Hongguang Ma, Felipe Carvalho"
    affiliation: "Pacific Islands Fisheries Science Center"

# Table of contents and navigation
toc: true
toc-title: "Table of contents"
toc-location: right
toc-depth: 2
number-sections: true

filters:
  - fontawesome 

format:
  html:
    toc-float: true
    embed-resources: true

execute: 
  echo: false
---

```{r}
#| label: setup
#| include: false
#| message: false
#| warning: false
library(tidyverse)
library(gt)
# renv::install("katiejolly/nationalparkcolors")
library(nationalparkcolors)
arches_colors <- nationalparkcolors::park_palette(name = "Arches", n =5)

```

[{{< fa brands github >}} GitHub Repository](https://github.com/PIFSCstockassessments/Opaka_simulation){target="_blank" .btn .btn-light}

# Introduction  
 
## Background and Context

*  Overview of Opakapaka (Pristipomoides filamentosus) in the Main Hawaiian Islands  
*  Ecological and economic importance of the fishery
*  Current management framework and challenges  
*  Data limitations and uncertainties in current assessments
*  Recruitment variability and its impact on population dynamics
*  Challenges specific to deep-water snappers in tropical systems

## Survey Design and Optimization

*  The need for efficient survey design given resource constraints
*  Trade-offs between survey effort and assessment precision
*  Importance of understanding survey performance under different scenarios

## Future Uncertainty and Recruitment Scenarios

*  Climate change and environmental variability impacts on recruitment
*  The need to evaluate survey performance under contrasting future conditions
*  Why poor recruitment scenarios are particularly important to consider

 The objective of this project is to understand how we optimize an operational fishery-independent survery for Opakapaka in the Main Hawaiian Islands by testing the impact of effort of the survey under two future recruitment conditions on our ability to estimate key stock quantities such as spawning stock biomass and management reference points. 

# Methods 

## Operating Model  

The operating model was created using a Stock Synthesis (#TODO: ADD CITATON - METHOT and WETZEL) model, conditioned on life history of the Opakapaka stock and trends of historical data available for the Deep 7 fishery. Catch, CPUE, and size data from the commerical and non-commercial fisheries were available for 75 years and index of abundance and length composition data from the Bottomfish Fishery-Independent Survey in Hawaii (BFISH) for seven years. The OM was used to simulate the population dynamics of the stock and data that were sampled with observation error and fit in the estimation model (EM) for 100 years. To generate data for the EM, we ran the OM with a time series fishing mortality and recruitment deviations. We generated 100 time series of fishing mortality based on the trend found in the historical data, low at the start of the fishery, ramp up in the 1970s and remain high until the late 1990s, then decline in the early 2000s and remain constant until the end of the time series (@fig-F-timeseries). Some variability (standard deviation of 0.002) was introduced to the annual fishing mortality values so that there was some sotchasticity between iterations. Following the approach in the most recent benchmark stock assessment for the Deep 7 fishery, we used a time-varying ratio for the non-commercial catch (#TODO: add citation and ratio values). 

```{r}
#| message: false
#| label: fig-F-timeseries
#| fig-cap: Fishing mortality time series for the commercial (blue lines) and non-commercial (orange lines) fisheries.
load("../Inputs/constantF_mat.RData")

comm <- F_comm_df %>% 
as.data.frame() %>% 
mutate(Year = seq(1949, 2048), 
Fishery = "Commercial") %>%
pivot_longer(cols = -c("Year", "Fishery"), values_to = "Fishing Mortality") %>%
mutate(number = as.numeric(str_remove(name, "V"))) 

noncomm <- F_noncomm_df %>% 
as.data.frame() %>% 
mutate(Year = seq(1949, 2048), 
Fishery = "Non-Commercial") %>%
pivot_longer(cols = -c("Year", "Fishery"), values_to = "Fishing Mortality") %>%
mutate(number = as.numeric(str_remove(name, "V"))) 

comm %>% 
bind_rows(noncomm) %>%
ggplot(aes(x = Year, y = `Fishing Mortality`)) + 
geom_line(aes(group = interaction(number, Fishery), color = Fishery), alpha = .65) +
stockplotr::theme_noaa() +
  scale_colour_manual(
  values = nationalparkcolors::park_palette(name = "Arches", n=2), 
  name = "Fishery"
  ) 

```

The last 25 years are considered "future years" because they represent possible future recruitment conditions that we want to test the EM's robustness against. We tested two future recruitment conditions, the first was a "normal" recruitment, assuming no changes in mean recruitment over time and the second was a "poor recruitment" scenario, which assumed a decline in mean recruitment by 40% in the future. 

### Recruitment Deviation Time Series Generation

We developed 200 stochastic recruitment deviation time series for use in the operating models to evaluate fishery management performance under different recruitment patterns. Two distinct recruitment scenarios were created: (1) normal future recruitment and (2) poor future recruitment, represented by a 40% decline in mean expected recruitment. Both scenarios had recruitment deviations generated from a standard normal distribution in log space for the first 75 years. 

Deviations were bias corrected as follows:
$$
R_{dev}(y,i) = σ_R × ε(y,i) - \frac{{σ_R}^2}{2}
$$


where $R_{dev}(y,i)$ is the recruitment deviation for year y and iteration i, $ε(y,i) \sim N(0,1)$, and the bias correction term ${σ_R}^2$ ensures unbiased recruitment in expectation. $σ_R$ was 0.52 (#TODO: LH citation). 

### Future Recruitment Condition

#### Normal Recruitment 

For the normal recruitment condition, 100 future recruitment deviations were generated assuming recruitment would continue following the same stochastic process as the final assessment years:

$$
R_{dev}(y,i) = σ_R × ε(y,i) - \frac{{σ_R}^2}{2}
$$

where $σ_R$ = 0.15, maintaining consistency with the variability observed in recent years.

#### Declining Recruitment   

To evaluate the robustness of management strategies under conditions of sustained recruitment failure, we developed a "poor recruitment" scenario characterized by a prolonged decline in recruitment success followed by a period of persistently low recruitment levels.

The recruitment decline scenario was structured as a two-phase process spanning 25 years. The first phase consisted of a 10-year gradual decline period, during which recruitment decreased from baseline levels to 60% of the historical mean. The annual decline rate (*r*) was calculated as:
$$
r = (r_{reduced}^(\frac{1}{10})) - 1
$$
where *r~reduced~* is the proportion by which the stock was reduced. Because of the non-linear relationship between spawning biomass and recruitment, the reduced level did not actually need to be 60%. We used an iterative process to determine what reduced proportion actually resulted in 60% of the mean recruitment (*r~reduced~* = 0.71). 
This formulation ensured a smooth exponential decline reaching the target reduction after 10 years. The second phase maintained recruitment at the reduced level (60% of baseline) for an additional 15 years, representing a new regime of lower recruitment conditions.

The baseline recruitment level was established using the mean recruitment from the most recent 15-year period (279.67 individuals), representing contemporary recruitment conditions prior to the simulated decline. 

For the decline phase, annual multipliers were calculated as:
$$
M_t = (1 + r)^t
$$

where *t* represents the ten years of the decline. For the lower recruitment regime period (years 11-25), multipliers were set to *r~reduced~*, maintaining recruitment at 60% of baseline levels.

Projected recruitment levels were converted to recruitment deviations in log-space to maintain consistency with stock assessment model parameterization. Base recruitment deviations were calculated as:

$$
δ_t = ln(\frac{R_t}{R̄}) - \frac{{σ_R}^2}{2}
$$

where *R~t~* represents projected recruitment in year *t*, *R̄* is the baseline mean recruitment (279.67), and $σ_R$ is the recruitment variability parameter (0.52). The bias correction term $(σ_R^2 / 2)$ accounts for the Jensen's inequality effect when transforming between arithmetic and log scales.

To incorporate realistic recruitment variability around the deterministic decline trend, we generated 100 Monte Carlo iterations of the poor recruitment scenario. For each iteration, recruitment deviations were modified by adding random noise drawn from a normal distribution with mean = 0 and standard deviation = 0.15. This additional variability was intentionally set smaller than the base recruitment variability ($σ_R$ = 0.52) to ensure the underlying decline signal remained detectable above the noise.

The final recruitment deviations for each iteration *i* were calculated as:
$$
δ_{t,i} = δ_t + ε_{t,i}
$$

where $ε_{t,i} \sim N(0, 0.15²)$.

```{r}
#| message: false
#| label: fig-rec_dev
#| fig-cap: Log recruitment deviations for one iteration used in the normal (blue line) and poor (orange line) recruitment scenarios.
 ts_df <- read.csv("../all_scens_ts.csv")

ts_df %>%
filter(model_run == "om" & str_detect(scenario, "IRF_") & iteration == 14) %>%
mutate(scenario = factor(scenario, levels = c("IRF_SQ_25_yrfwd", "IRF_poorrec_25_yrfwd"))) %>% 
ggplot() + 
geom_hline(yintercept = 0, linewidth = 1.2, color = "grey60") +
geom_line(aes(x = year, y = rec_dev, group = scenario, color = scenario), linewidth = 1.3) +
labs(x = "Year", y = "Log Recruitment Deviation") + 
stockplotr::theme_noaa() +
scale_colour_manual(
  values = nationalparkcolors::park_palette(name = "Arches", n=2), 
  name = "Recruitment Scenario", labels = c("Normal", "Poor")
  ) 

```

### Life History

Growth was modeled using a von Bertanlanfy growth curve and natural mortality was assumed to be 0.135 and age-invariant. Weight-at-length relationship, maturity and fecundity. The stock-recruitment relationship was modeled with a Beverton-Holt relationship where steepness was fixed at 0.76 and $σ_R$ was 0.52.  

### Initial Fishing Mortality and Selectivity

Initial fishing mortality was fixed at 0.012 for the commerical fishery and assumed to start at equilibrium for the non-commercial fishery. Logistic selectivity was modeled for the commercial and non-commercial fisheries and a dome-shaped double normal selectivity model was used for the research fishing survey (@fig-selex). 

```{r}
#| warning: false
#| message: false
#| label: fig-selex
#| fig-caption: "Selectivity functions used for the commercial (blue line), non-commercial (orange line), and survey (red line)."
rep <- r4ss::SS_output(dir = "../models/02_Opaka_BFISH_simulation_EM", verbose = F, printstats=FALSE)

rep$sizeselex %>% 
filter(Yr == 2023 & Factor == "Lsel") %>%
select(-c("Factor", "Yr", "Sex", "Label")) %>%
pivot_longer(cols = -Fleet, names_to = "Length", values_to = "Selectivity") %>%
mutate(Length = as.numeric(Length),
Fleet = case_when(
  Fleet == 1 ~ "Commercial",
  Fleet == 2 ~ "Non Commercial",
  Fleet == 3 ~ "Survey"
)) %>%
ggplot(aes(x = Length, y = Selectivity)) +
geom_line(aes(color = Fleet), linewidth = 1.5) + 
stockplotr::theme_noaa() +
scale_colour_manual(values = nationalparkcolors::park_palette(name = "Arches", n=3)) 
```


### Data Generation  

We evaluated five sampling scenarios to assess how different levels of survey effort affect stock assessment performance under each recruitment condition (@tbl-sas). Survey effort was defined by the number of sampling grids covered, which directly influenced the precision of the resulting abundance index and length composition data. Catch, CPUE, and length composition data were simulated from the OM based on the population dynamics generated by the specific F and recruitment deviation time series (@fig-data). 

**Survey Effort Scenarios**  
The scenarios ranged from high survey effort (HSE) with 632 grids to extra low survey effort (XLSE) with only 257 grids. As expected, higher effort translated to more precise data: index CVs ranged from 15% (HSE) to 30% (XLSE), and effective sample sizes for length compositions ranged from 60 (HSE) to 15 (XLSE). The no survey effort scenario (NSE) represented a baseline case where stock assessments would rely solely on fishery-dependent data.

```{r}
#| label: tbl-sas
#| tbl-cap: "The observational error used to sample data from the OM for high survey effort (HSE), intermediate survey effort (ISE), low survey effort (LSE), extra low survey effort (XLSE), and no survey effort (NSE). For each survey effort, the number of sampling grids (Grids) that effort would translate to, the resulting index of abundance CV, and effective sample size for length composition data are shown. All sampling effort scenarios were tested under both normal and poor recruitment conditions."
 
sas <- read.csv("sas.csv")

CV <- c(.24, .18, .15, .3)
# equation based on fitting linear regression to Model-based CV for resfish ~ # resfish grids for years 2018-2022
x <- (CV - 0.4027)/-0.0004

sas_table <- sas %>% 
select(-c(N_years, F_scen, Resfish_sd_obs, Recruitment)) %>%
mutate(
    Scen_name = str_remove(Scen_name, "_SQ"),
    Scen_name = str_remove(Scen_name, "_poorrec"),
    N_grids = c(0,0,407,407,557,557,632,632,557,557,257,257)) %>% 
    filter(str_detect(Scen_name, "Selex", negate = TRUE)) %>%
distinct(Scen_name, .keep_all = T) %>%
select(Scen_name, N_grids, Resfish_index_CV, Neff_len_Resfish) 

knitr::kable(sas_table,
             col.names = c("Sampling Effort", "Grids", "CV", "effN"),
             align = "lccc") %>%
  kableExtra::kable_styling(bootstrap_options = c("condensed", "hover"),
                           full_width = FALSE,
                           position = "center",
                           font_size = 12)
```

**Data Sources by Scenario**  
All scenarios had some combination of catch, abundance index, and length composition data available for the EM (@fig-data). For all survey scenarios (HSE through XLSE), commercial CPUE and length data were discontinued after 2023, forcing the estimation model to rely entirely on survey-based abundance and length data for the final 25 years. This design allowed us to isolate the effect of survey data quality on assessment performance. In contrast, the NSE scenario continued using commercial fishery data throughout the entire time series.

```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 12
#| label: fig-data
#| fig-cap: "Data generated from operating models and used in the estimation models with the survey (a) and the estimation models without the survey (b). The size of each point represents the relative amount of data available per year and fleet (i.e. large points represent years with more data)."
#include figure to show type/amount of data for each scenario
reps <- r4ss::SSgetoutput(
  dirvec = c("../FRSonly_SQ_25_yrfwd/1/em",
  "../HRF_SQ_25_yrfwd/1/em"))

# Create custom layout with larger plots
layout(matrix(c(2,1), nrow = 2), widths = c(1), heights = c(2,2))
par(mar = c(5,4,4,2))

# First plot
r4ss::SSplotData(reps$replist1, subplots = 2, 
                 fleetcol = c(arches_colors[1], arches_colors[2]), 
                 fleetnames = c("Commercial", "Non Commercial"))
title("b) Without survey", adj = 0, font.main = 2)

# Second plot
r4ss::SSplotData(reps$replist2, subplots = 2, 
                 fleetcol = c(arches_colors[1], arches_colors[2], arches_colors[3]), 
                 fleetnames = c("Commercial", "Non Commercial", "Survey"))
title("a) With survey", adj = 0, font.main = 2)
```


**Hyperstability in Fishery Data**  
For the NSE scenario under poor recruitment conditions, we incorporated hyperstability into the commercial CPUE data to reflect realistic fishery behavior during population decline. The CPUE and catch data were modified to not fully reflect the underlying population decline, simulating the tendency of commercial fisheries to maintain catch rates even as fish abundance decreases through improved fishing efficiency, targeting behavior, or fishing in higher-density areas.

**Observation Error Specification**  
Sampling error (variability in data sampled from the OM) and observation error specified in the EM were the same, ensuring no misspecification of uncertainty. The CV and effective sample size parameters remained constant across all years within each scenario (@tbl-sas).


## Estimation Model  
Data generated from the OM were fit using a catch-at-age model in Stock Synthesis. All life history and fishery relationships were correctly specified as in the OM. The only difference in the EM from the OM was for the poor recruitment scenarios, a regime time block was added for the last 15 years. Without the regime, the model was unable to fit the persistant decline in recruitment and the relative errors in outputs were over exaggerated. When fitting the model, all parameters were fixed except for R~0~, the regime parameter for the poor recruitment scenarios, catchability for the commercial and survey fleets, and selectivity parameters for the commercial fishery and survey fleet.

In total, 1,000 simulations were run using the Open Science Grid #TODO: add citation. Each simulation run involved generating the population dynamics based on the OM, with the given F and recruitment deviations timeseries, sampling data with error from the OM, and fitting those data in an EM. This process was done using the r package {ss3sim} #TODO add citation. 

## Performance Indices  
To compare the bias and precision of each sampling strategy, we looked at median relative error (MRE) and median absolute relative error (MARE) between the OM and EM for spawning stock biomass (SSB), fishing mortality (F), age-0 recruitment, and reference points SSB/SSB~MSST~ and F/F~MSY~. 

$$
MRE = median(\frac{EM - OM}{OM})
$$

$$ 
MARE = median[abs(\frac{EM - OM}{OM})]
$$

where EM is the estimated value from the EM and OM is the true value from the OM. 

# Results  

:::{.callout-note icon=false}
## Key results 

### Normal Recruitment Conditions  {.unnumbered}

* All sampling strategies (including no survey effort) showed similar bias for SSB and F estimates (MRE 2-3.6% for SSB, -2.4% to -3.3% for F)  
* While bias remained similar, uncertainty (error ranges) increased substantially as sampling effort decreased, with XLSE showing -20% to +35% relative error for SSB  
* Terminal year F and F/FMSY estimates showed greater sensitivity to extremely low sampling effort (XLSE) compared to SSB-based metrics  
* Persistent overestimation of recruitment led to optimistic bias in SSB estimates across all scenarios

### Poor Recruitment Conditions  {.unnumbered}

* No survey effort scenario showed dramatically higher error rates (42% MARE for SSB) compared to other scenarios (9-14% MARE)  
* Performance gaps between high and low effort scenarios widened substantially under recruitment failure  
* All models failed to accurately capture poor stock conditions, with most runs showing overly optimistic stock status estimates (interquartile ranges rarely included true values)  
* While NSE captured true F/FMSY values more often, it came with much wider uncertainty ranges, potentially leading to management overreactions  

:::

Out of 1,000 EM runs, 8 runs did not converge (maximum gradient was greater than 1e-4). The EM with NSE under poor recruitment conditions had the lowest convergence rate with four iterations that did not converge but overall, model convergence for all sampling scenarios was high. 

## Normal Recruitment Conditions 

We found that under normal recrutiment conditions, all sampling strategies were similiarly biased (@fig-mre) and precise (@fig-ssbmare) for SSB. MRE for the terminal year ranged from 2% to 3.6%, and MARE for the terminal year ranged from 5.8% to 10.6%. Even the scenario where there was no survey effort and just the commercial fishery CPUE and length data, the MRE was slightly smaller than the highest effort scenario. This was likely due to the length of the CPUE time series and the effective sample sizes of the length data from the fishery. Uncertainty around SSB, particularly in the last 25 years increased with decreasing sampling effort. The XLSE had the largest uncertainty, ranging between -20% and 35% relative error from the true values. This discrepency can have large impacts on reference point estimation.   


```{r}
ts_re <- read.csv("ts_re.csv")
```

```{r}
#| label: fig-mre
#| message: false
#| warning: false
#| fig-cap: "Timeseries of median (line) and 95% interval (band) relative error of spawing stock biomass for the high (HSE), intermediate (ISE), low (LSE), extra low (XLSE), and no survey effort (NSE) scenarios under both the normal and poor recruitment conditions."

ts_re %>%
filter(year <= 2048 & str_detect(scenario, "Selex", negate = T)) %>%
group_by(scenario, year) %>% 
summarise(med_re = quantile((SpawnBio_re), 0.5, na.rm = TRUE),
uci = quantile((SpawnBio_re), .975, na.rm = TRUE),
lci = quantile((SpawnBio_re), .05, na.rm = TRUE)) %>% 
separate(col = scenario, into = c("scen", "Rec_scenario", "25", "Yrs"), sep = "_",
remove = F) %>% 
mutate(Rec_scenario = ifelse(Rec_scenario == "poorrec", "Poor Recruitment", "Normal Recruitment")) %>%
ggplot() +
geom_hline(yintercept = 0, color = "grey60") +
geom_ribbon(aes(x = year, ymin = lci, ymax = uci, fill = scen, color = scen), 
alpha = .25, linewidth = .5) + 
geom_line(aes(x = year, y = med_re, color = scen), linewidth = 1.2) +
facet_wrap(~ Rec_scenario) + 
labs(x = "Year", y = "SSB Relative Error") + 
stockplotr::theme_noaa() +
  scale_colour_manual(
    values = arches_colors,
    name = "Sampling Scenario", 
    labels = c("NSE", "HSE", "ISE", "LSE", "XLSE")
  ) +
  scale_fill_manual(
    values = arches_colors,
    name = "Sampling Scenario", 
    labels = c("NSE", "HSE", "ISE", "LSE", "XLSE")
  )

```


```{r}
#| label: fig-ssbmare
#| message: false
#| warning: false
#| fig-cap: "Timeseries of median (line) and 95% interval (band) absolute relative error of spawing stock biomass for the high (HSE), intermediate (ISE), low (LSE), extra low (XLSE), and no survey effort (NSE) scenarios under both the normal and poor recruitment conditions."

ts_re %>%
filter(year <= 2048 & str_detect(scenario, "Selex", negate = T)) %>%
group_by(scenario, year) %>% 
summarise(med_re = quantile(abs(SpawnBio_re), 0.5, na.rm = TRUE),
uci = quantile(abs(SpawnBio_re), .975, na.rm = TRUE),
lci = quantile(abs(SpawnBio_re), .05, na.rm = TRUE)) %>% 
separate(col = scenario, into = c("scen", "Rec_scenario", "25", "Yrs"), sep = "_",
remove = F) %>% 
mutate(Rec_scenario = ifelse(Rec_scenario == "poorrec", "Poor Recruitment", "Normal Recruitment")) %>%
ggplot() +
geom_hline(yintercept = 0, color = "grey60") +
geom_ribbon(aes(x = year, ymin = lci, ymax = uci, fill = scen, color = scen), 
alpha = .25, linewidth = .5) + 
geom_line(aes(x = year, y = med_re, color = scen), linewidth = 1.2) +
facet_wrap(~ Rec_scenario) + 
labs(x = "Year", y = "SSB Absolute Relative Error") + 
stockplotr::theme_noaa() +
  scale_colour_manual(
    values = arches_colors,
    name = "Sampling Scenario", 
    labels = c("NSE", "HSE", "ISE", "LSE", "XLSE")
  ) +
  scale_fill_manual(
    values = arches_colors,
    name = "Sampling Scenario", 
    labels = c("NSE", "HSE", "ISE", "LSE", "XLSE")
  )

```

We found that fishing mortality also had similar bias (@fig-Fmre) and precision (@fig-Fmare) for all scenarios, however the bias was negative (estimates were slightly under estimated when compared to the true values). Terminal year F MRE ranged between -2.4% and -3.3%. While MRE for all scenarios was similar for SSB and F, the uncertainty increased as the survey effort decreased, except for the NSE scenario which had similar uncertainty to the HSE scenario. The uncertainty also increased for all scenarios except the NSE scenario in the future years. Again, this was likely due to the exclusion of fishery CPUE and length data for those years and the short overlap of time between the survey and fishery data. 

```{r}
#| label: fig-Fmre
#| message: false
#| warning: false
#| fig-cap: "Timeseries of median (line) and 95% interval (band) relative error of fishing mortality for the high (HSE), intermediate (ISE), low (LSE), extra low (XLSE), and no survey effort (NSE) scenarios under both the normal and poor recruitment conditions."

ts_re %>%
filter(year <= 2048 & str_detect(scenario, "Selex", negate = T)) %>%
group_by(scenario, year) %>% 
summarise(med_re = quantile((F_1_re), 0.5, na.rm = TRUE),
uci = quantile((F_1_re), .975, na.rm = TRUE),
lci = quantile((F_1_re), .05, na.rm = TRUE)) %>% 
separate(col = scenario, into = c("scen", "Rec_scenario", "25", "Yrs"), sep = "_",
remove = F) %>% 
mutate(Rec_scenario = ifelse(Rec_scenario == "poorrec", "Poor Recruitment", "Normal Recruitment")) %>%
ggplot() +
geom_hline(yintercept = 0, color = "grey60") +
geom_ribbon(aes(x = year, ymin = lci, ymax = uci, fill = scen, color = scen), 
alpha = .25, linewidth = .5) + 
geom_line(aes(x = year, y = med_re, color = scen), linewidth = 1.2) +
facet_wrap(~ Rec_scenario) + 
labs(x = "Year", y = "F Relative Error") + 
stockplotr::theme_noaa() +
  scale_colour_manual(
    values = arches_colors,
    name = "Sampling Scenario", 
    labels = c("NSE", "HSE", "ISE", "LSE", "XLSE")
  ) +
  scale_fill_manual(
    values = arches_colors,
    name = "Sampling Scenario", 
    labels = c("NSE", "HSE", "ISE", "LSE", "XLSE")
  )

```

```{r}
#| label: fig-Fmare
#| message: false
#| warning: false
#| fig-cap: "Timeseries of median (line) and 95% interval (band) absolute relative error of fishing mortality for the high (HSE), intermediate (ISE), low (LSE), extra low (XLSE), and no survey effort (NSE) scenarios under both the normal and poor recruitment conditions."

ts_re %>%
filter(year <= 2048 & str_detect(scenario, "Selex", negate = T)) %>%
group_by(scenario, year) %>% 
summarise(med_re = quantile(abs(F_1_re), 0.5, na.rm = TRUE),
uci = quantile(abs(F_1_re), .975, na.rm = TRUE),
lci = quantile(abs(F_1_re), .05, na.rm = TRUE)) %>% 
separate(col = scenario, into = c("scen", "Rec_scenario", "25", "Yrs"), sep = "_",
remove = F) %>% 
mutate(Rec_scenario = ifelse(Rec_scenario == "poorrec", "Poor Recruitment", "Normal Recruitment")) %>%
ggplot() +
geom_hline(yintercept = 0, color = "grey60") +
geom_ribbon(aes(x = year, ymin = lci, ymax = uci, fill = scen, color = scen), 
alpha = .25, linewidth = .5) + 
geom_line(aes(x = year, y = med_re, color = scen), linewidth = 1.2) +
facet_wrap(~ Rec_scenario) + 
labs(x = "Year", y = "F Absolute Relative Error") + 
stockplotr::theme_noaa() +
  scale_colour_manual(
    values = arches_colors,
    name = "Sampling Scenario", 
    labels = c("NSE", "HSE", "ISE", "LSE", "XLSE")
  ) +
  scale_fill_manual(
    values = arches_colors,
    name = "Sampling Scenario", 
    labels = c("NSE", "HSE", "ISE", "LSE", "XLSE")
  )

```


We found that the HSE and ISE best estimated the true age-0 recruitment of the population over the time series and the XLSE estimated the true age-0 recruitment the worst, particularly in the last few years (@fig-recmare). The persistent over estimation of age-0 recruitment for all models helps to explain the positive bias in SSB over time, particularly in the last few years of the simulation.  

```{r}
#| label: fig-recmare
#| message: false
#| warning: false
#| fig-cap: "Timeseries of median (line) absolute relative error of age-0 recruitment for the high (HSE), intermediate (ISE), low (LSE), extra low (XLSE), and no survey effort (NSE) scenarios under both the normal and poor recruitment conditions."

ts_re %>%
filter(year <= 2047 & str_detect(scenario, "Selex", negate = T)) %>%
group_by(scenario, year) %>% 
summarise(med_re = quantile(abs(Recruit_0_re), 0.5, na.rm = TRUE),
uci = quantile(abs(Recruit_0_re), .975, na.rm = TRUE),
lci = quantile(abs(Recruit_0_re), .05, na.rm = TRUE)) %>% 
separate(col = scenario, into = c("scen", "Rec_scenario", "25", "Yrs"), sep = "_",
remove = F) %>% 
mutate(Rec_scenario = ifelse(Rec_scenario == "poorrec", "Poor Recruitment", "Normal Recruitment")) %>%
ggplot() +
geom_hline(yintercept = 0, color = "grey60") +
geom_line(aes(x = year, y = med_re, color = scen), linewidth = 1.2) +
facet_wrap(~ Rec_scenario) + 
labs(x = "Year", y = "Absolute Relative Error of Age-0 Recruitment") + 
stockplotr::theme_noaa() +
  scale_colour_manual(
    values = arches_colors,
    name = "Sampling Scenario", 
    labels = c("NSE", "HSE", "ISE", "LSE", "XLSE")
  ) +
  scale_fill_manual(
    values = arches_colors,
    name = "Sampling Scenario", 
    labels = c("NSE", "HSE", "ISE", "LSE", "XLSE")
  )
```

Under normal recruitment conditions, all models performed similarly when estimating the stock status (SSB/SSB~MSST~ and F/F~MSY~). The median relative error of SSB/SSB~MSST~ was slightly over-estimated compared to the true status, less than 10% for all model runs. However the range of relative error increased as sampling effort decreased, with the exception of the NSE scenario.  The median relative error for all models was slightly underestimated for F/F~MSY~ and the range increased with decreasing sampling effort. We found that terminal year F and F/F~MSY~ is more sensitive to XLSE than terminal year SSB and SSB/SSB~MSST~. However, the other sampling scenarios seem to be robust to decreasing sampling effort for F and F~MSY~.  

```{r}
#| message: false
#| warning: false
#| label: fig-refpoints
#| fig-cap: Distirbution of relative error for terminal year SSB, spawning stock biomass at minimum stock size threshold (SSB_MSST), the terminal year stock status (SSB/SSB_MSST), terminal year F, FMSY, and terminal year fishing status (F/FMSY). 

ref_pt_re <- read.csv("ref_pt_re.csv")

ref_pt_re %>%
    select(c(iteration, scenario, SpawnBio_re, ssb_ssbmsst_re, annual_F_re, F_Fmsy_re)) %>%
    filter(scenario != "IRFSelex_poorrec_25_yrfwd") %>%
    rename(
        "iteration" = iteration,
        "scenario" = scenario,
        "Terminal~Year~SSB" = SpawnBio_re,
        "SSB/SSB[MSST]" = ssb_ssbmsst_re,
        "Terminal~Year~F" = annual_F_re,
        "F/F[MSY]" = F_Fmsy_re
    ) %>%
    pivot_longer(cols = -c(iteration, scenario)) %>%
    separate(
        col = scenario, into = c("scen", "Rec_scenario", "25", "Yrs"), sep = "_",
        remove = F
    ) %>%
    mutate(
        name = factor(name,
            levels = c(
                "Terminal~Year~SSB", "SSB/SSB[MSST]",
                "Terminal~Year~F", "F/F[MSY]"
            )
        ),
        Rec_scenario = ifelse(Rec_scenario == "poorrec", "Poor~Recruitment", "Normal~Recruitment")
    ) %>%
    ggplot() +
    geom_boxplot(aes(x = scen, y = value, fill = scen)) +
    geom_hline(yintercept = 0, color = "grey50", linetype = 2) +
    facet_grid(Rec_scenario ~ name, scales = "free_y", labeller = label_parsed) +
    scale_x_discrete(labels = NULL, breaks = NULL) +
    labs(x = "", y = "Relative Error") +
    stockplotr::theme_noaa() +
    scale_fill_manual(
        values = arches_colors,
        name = "Sampling Scenario",
        labels = c("NSE", "HSE", "ISE", "LSE", "XLSE")
    )


```

## Poor Recruitment Conditions 

 When tested under poor recruitment conditions, we found that most models performed similarly in terms of bias and precision for SSB (@fig-mre @fig-ssbmare) and F (@fig-Fmre @fig-Fmare), however the MRE and MARE was greater for all models than under normal recruitment scenarios. Terminal SSB MARE ranged from 9.3% to 14% for all models except the NSE. NSE had a much higher error, 42% MARE for SSB. Terminal F MARE ranged from 9.3% to 13% for all models.The MARE for NSE showed an interesting pattern compared to the rest of the sampling scenarios for the F timeseries. MARE started off much higher than the other scenarios but during the period of poor recruitment, the MARE slightly decreased and was similar to the other sampling scenarios. This could be due to _____ #TODO  

We found that under poor recruitment conditions, HSE, ISE, and LSE perform similarly to the HSE and ISE models under normal recruitment scenarios (@fig-recmare). However, XLSE performs slightly worse and NSE performs the worst, with MARE increasing to over 40% in the last few years of the model. 

When evaluating the performance of estimating stock status reference points, we found that the HSE and ISE scenarios performs the best for terminal year SSB and SSB/SSB~MSSST~ (@fig-refpoints). The median relative errors are closest to 0 for both quantites and the range of uncertainty is the smallest. NSE performs the worst, with the median relative error around 40% for both terminal year SSB and SSB/SSB~MSST~. The range of uncertainty becomes notably wider across sampling scenarios, with the largest range under the NSE scenario. It should be noted though that the true terminal year SSB or SSB/SSB~MSST~ is not recovered for the majority of runs under any scenario (the interquartile range does not include 0). This suggests that under poor recruitment conditions, all of the tested models would provide an overly optimistic view of the stock status. A similar pattern held true for terminal year F and F/F~MSY~ HSE had the smallest range of error and the range become wider as sampling effort decreased, except for the NSE scenario. For fishing mortality based estimates, the NSE scenario performed well. The MRE for both F and F/F~MSY~ was close to 0, and the true F and F/F~MSY~ values were captured within 50% of the runs (0 is included in the interquartile range). However, while that scenario does capture the true value more closely, the trade-off is that the range of error is much wider than any of the other scenarios. This could lead to a greater over estimation of F/F~MSY~ which could have negative impacts on the fishery such as unnecessary reduction in allocations or closures. 


# Discussion  

:::{.callout-note icon=false}

## Key points  

* The poor recruitment condition reveals the importance of adequate sampling  
* Management reference points become more uncertain as sampling effort decreases  
* The NSE scenario suggests that it would be adequate in some situations, however, there are a few critical assumptions that would need to be considered before taking this approach  
* During periods of recruitment failure, enhanced rather than reduced monitoring may be necessary to maintain assessment reliability  
* Managers should account for increased uncertainty in stock status during periods of both poor recruitment and reduced survey effort  
* While reduced survey effort may seem economically attractive, the substantial increases in assessment uncertainty could lead to suboptimal management decisions with significant long-term costs

:::

In this simulation, we found that if normal recruitment persists in the future, the sampling effort of the fishery independent survey does not impact our ability to estimate the biomass, fishing mortality, and stock status in general, but it does increase our uncertainty around those estimates. On the other hand, we found that if recruitment were to decline in the future and a regime-type shift occurred, the survey effort impacts our ability to estimate biomass, fishing mortality, and stock status and the uncertainty around those estimates as well. Particularly, if there is no fishery-independent survey, our ability to estimate key quantities deteriorates significantly.

The poor recruitment scenario revealed critical thresholds for maintaining assessment reliability under challenging population conditions. High and intermediate survey effort scenarios (HSE and ISE) demonstrated comparable performance across all metrics, including median relative error (MRE), median absolute relative error (MARE), and the 95% confidence intervals of relative error. Lower effort scenarios (LSE and XLSE) maintained similar central tendency measures but exhibited substantially wider confidence intervals, indicating reduced precision in parameter estimates. This pattern illustrates how poor recruitment conditions amplify the negative consequences of reduced sampling effort, creating a compounding effect where assessment reliability deteriorates precisely when accurate population monitoring becomes most critical for management decisions. The performance degradation observed below intermediate effort levels suggests that survey coverage should be maintained between 550-630 grids to ensure adequate assessment quality for Opakapaka stock management, particularly given the potential for unexpected recruitment variability in this system.

One interesting finding was the stark contrast of the NSE scenario performances under the two recruitment conditions. The NSE scenario's performance suggests that in some cases, no survey data may be adequate if the fishery CPUE index is accurately tracking the population trends. Having one continuous data source of good quality may be preferable to shorter time series of data or very poor quality survey data. However, this is only true under the normal recruitment conditions and once poor recruitment was introduced the model performed far worse than the others in almost every metric. A major assumption of the NSE model is that the fishery data, particularly the CPUE and size data are accurately characterizing the full population. This is likely untrue, as fishers are knowledgeable about where and when to find larger fish and will likely be able to maintain a healthy catch even if the population is declining. Therefore, we should not assume that the fishery is an accurate, unbaised representation of what is actually happening in the population.  

Our simulation relies on several key assumptions that may not fully capture the complexity of real-world fishery dynamics and data collection challenges. The NSE scenario under poor recruitment assumed complete hyperstability in fishery data, meaning catch rates remained unaffected by population decline throughout the 25-year projection period. While hyperstability is a documented phenomenon in fishery-dependent CPUE indices, sustained recruitment failure would likely eventually manifest in commercial catch data, suggesting our model may have overestimated the NSE scenario's limitations. Nevertheless, this represents a plausible worst-case scenario that highlights the risks of relying solely on fishery-dependent data during periods of stock decline. Additionally, our model simplified future data availability by excluding fishery CPUE and length composition data after 2023, which artificially increased uncertainty even in high survey effort scenarios. In practice, these fishery data streams would continue but would be appropriately weighted within the assessment framework to balance their influence against survey observations. Finally, the poor recruitment scenarios employed reduced variability in recruitment deviations to clearly establish declining trends, whereas natural recruitment variability typically remains high even during periods of average decline. This simplification may have amplified the apparent impacts of recruitment failure, as realistic variability could obscure declining trends and complicate early detection of regime shifts.

While consistent long-term monitoring remains ideal, a combination of adaptive sampling strategies and improved data collection techniques could help balance assessment reliability with resource constraints. An adaptive approach could involve reducing survey effort during periods when recruitment appears stable and population indicators show no signs of decline, then increasing sampling intensity when early warning signals suggest recruitment failure or population stress. However, such adaptive strategies must be implemented cautiously, as maintaining data continuity over extended periods provides irreplaceable value for detecting long-term trends and regime shifts. Complementary improvements to data collection could enhance assessment quality without necessarily increasing survey effort, such as incorporating annual age data from the fishery-independent survey to better understand population structure and recruitment dynamics, or refining fishing methodology to reduce coefficient of variation in catch estimates. 

These results highlight several critical considerations for fishery management. The first is that maintaining adequate survey effort is essential for reliable stock assessments, particularly for fishing mortality estimates used in management decisions. The second is that during periods of recruitment failure, enhanced rather than reduced monitoring may be necessary to maintain assessment reliability. We saw a noticeable decrease in our estimation abilities when recruitment was poor and there was a reduced survey effort. Additionally, scientists and managers should account for increased uncertainty in stock status during periods of both poor recruitment and reduced survey effort (even under normal recruitment conditions). Lastly, while reduced survey effort may seem economically attractive, the substantial increases in assessment uncertainty could lead to suboptimal management decisions with significant long-term costs for the fishery.
 