---
title: "Creating the OMs"
format: html
---


```{r}
library(r4ss)
library(tidyverse)
main.dir <- this.path::here(.. = 1)
source(file.path(main.dir, "R", "simF.r"))
set.seed <- read.csv(file.path(main.dir, "Inputs", "setseed.csv"))
niter <- 100
```


### Create vectors of values for OMs
```{r}
#Model estimated Fs
rep <- SS_output(dir = file.path(main.dir, "models", "03-simplified-mod"))
F_comm_mod <- rep$exploitation$Comm[-1]
F_noncomm_mod <- rep$exploitation$Non_comm[-1]

#create descreasing F trajectories
# f_dec <- getFhist(nsim=10,Esd=0.1,nyears=75,dFmin=-0.2,dFmax=-0.1,bb=0.2,scale= 0.04)
# f_dec_rescale <- apply(f_dec, 1, FUN = function(x)(scales::rescale(x, to = c(0.00958269, 0.08221600)))) #put on a similar scale to real F
# png(filename = file.path(main.dir, "Model_Runs", "fid_example.png"),width=10,height=8,units="in",res=450,pointsize = 12)
# matplot(x=seq(1,75,1),(f_dec_rescale),type='l',ylim=c(0,0.1),xlab="",ylab="Fishing mortality", lwd = 3,cex.lab=1.5,cex.axis=1.5, col = 1:10)
# legend("topleft", legend = seq(1,10), col = 1:10, lty = 1, cex = 1.5)
# dev.off()

#create idc F trajectories
# log_sd <- 0.2
# sqrt(exp(log_sd**2)-1)
first_part <- sin_func(nyears = 60, low = 3*pi/2, high = 3*pi, Esd = 0.2)
second_part <- sin_func(nyears = 20, low = pi, amplitude = 0.5, Esd = 0.2)[2:16]
F_comm <- scales::rescale(c(first_part, second_part), to = c(0.0101208, 0.0868887)) #put on a similar scale to real F
# plot(x = seq(1,75), y = fsin, type = "l")
u_vec <- c(rep(1.8, 26), rep(1.47, 29), rep(1.03, 20))
F_noncomm <- u_vec * F_comm
#png(filename = file.path(main.dir, "Model_Runs", "fidi_example.png"),width=10,height=8,units="in",res=450,pointsize = 12)
plot(x = 1:75, y = F_comm, type = "l", ylim = c(0,0.15), ylab = "F", xlab = "Year")
lines(x = 1:75, y = F_noncomm, col = "blue")
lines(x = 1:75, y = F_comm_mod, lty = 2)
lines(x = 1:75, y = F_noncomm_mod, col = "blue", lty = 2)
legend("topleft", lty = c(1,1,2,2), col = c("black", "blue", "black", "blue"), legend = c("Commercial simulated", "Non-commercial simulated", "Commercial model", "Non-commerical model"))
#dev.off()

## creating multiple fsin vectors 
F_vec_mat <- matrix(NA, nrow = niter, ncol = 75)
for(i in 1:niter) {

first_part <- sin_func(nyears = 60, low = 3*pi/2, high = 3*pi, Esd = 0.2)
second_part <- sin_func(nyears = 20, low = pi, amplitude = 0.5, Esd = 0.2)[2:16]
F_comm <- scales::rescale(c(first_part, second_part), to = c(0.0101208, 0.0868887)) 

F_vec_mat[i,] <- F_comm

}
#png(filename = file.path(main.dir, "Model_Runs", "fidi_example.png"),width=10,height=8,units="in",res=450,pointsize = 12)
matplot(x = seq(1,75), y = t(F_vec_mat), type = "l", xlab = "Year", ylab = "F")
#dev.off()
#save(list = "F_vec_mat", file = file.path(main.dir, "Inputs", "Fsin_mat.RData"))
load(file.path(main.dir, "Inputs", "Fsin_mat.RData"))
```

### Create rec dev timeseries

```{r}

sd_vec = c(0.52) #value from SS model  
# set.seed(731)
rdevs_array <- matrix(NA, nrow =75, ncol = niter) ## yrs x reps x sds
for(r in 1:ncol(rdevs_array)){ 
  # create single rep of correlated vectors given sd
  rdev_faux  <- faux::rnorm_multi(n = nrow(rdevs_array), ## year vector
                                 vars = 1,
                                 mu = 0,
                                 r = 1-1e-5,
                                 sd = sd_vec)
  rdevs_array[,r] <- rdev_faux$X1
                                 
 
}
#png(file.path(main.dir, "Model_Runs", '2024-07-15-rdevs_array-100y.png'), height = 8, width =8, unit = 'in', res =400)
matplot(x = seq(1,75), y = rdevs_array, type = "l", lty = 1, lwd = 1.2)
abline(h = 0)
#dev.off()
round(mean(apply(rdevs_array,2,sd)),2) #check that SD makes sense
#save(list = "rdevs_array", file = file.path(main.dir, "Inputs", "recdevs_mat.RData"))
load(file.path(main.dir, "Inputs", "recdevs_mat.RData"))
```

### Create an OM and EM directories
```{r}
#| eval: false
iterations <- seq(1, niter, by = 1)
dir.it.om <- paste0(main.dir, "/Model_Runs/fidi/OM_", iterations)
sapply(dir.it.om, dir.create)
dir.it.em <- paste0(main.dir, "/Model_Runs/fidi/EM_", iterations)
sapply(dir.it.em, dir.create)

```

### Creating scenario variables

```{r}
#steepness
h_vec <- c(0.5, 0.76, 0.85)
#unreported catch ratios 
u_vec <- c(rep(1.8, 26), rep(1.47, 29), rep(1.03, 20))
##TODO: sampling vectors for catch, index, length and age comp data 
```

### Create starting OM and EMs

```{r}

for(I in 1:niter){
    # Get F-vector 
    F_noncomm <- u_vec * F_vec_mat[I,]
    F_list <- list(
        years = list(1:75, 1:75),
        fleets = c(1, 2), 
        fvals = list(F_vec_mat[I,], F_noncomm)
        )
    #create sampling scheme for indices of abundance
    index <- list(
        fleets = c(1,3), years = list(seq(1, 75, by = 1), seq(69, 75, by = 1)),
        seas = list(7,1), sds_obs = list(0.06, 0.25)
        )
    lcomp <- list(
        fleets = c(3), Nsamp = list(45),
        years = list(seq(69, 75, by = 1))
        )

    agecomp <- list(
        fleets = c(3), Nsamp = list(45),
        years = list(seq(69, 75, by = 1))
        )

    ss3sim_base(
        iterations = I,
        scenarios = "01-Fsin-mat-paka", 
        f_params = F_list,
        index_params = index,
        lcomp_params = lcomp,
        agecomp_params = agecomp,
        om_dir = om_dir,
        em_dir = em_dir,
        user_recdevs = as.matrix(rdevs_array[,I])
        )

}



```



### DEPRECATED: Loop to create OM and EMs
Maybe want to break this up so that creating the OMs first, then can create the EMs and sampling strategies. Especially for running on OSG, it will be better to have starting OMs and maybe EMs done already then can run the rest of simulation (feedback loop or just multiple scenarios)
```{r}
#| eval: false
file.list <- c("data.ss", "control.ss_new", "forecast.ss", "starter.ss", "ss_opt_win.exe")
for(i in 2:niter){
    om.dir <- dir.it.om[i]
    em.dir <- dir.it.em[i]

file.copy(file.path(main.dir, "models", "03-simplified-mod", "OM", file.list), om.dir, overwrite = TRUE) 
file.copy(file.path(main.dir, "models", "03-simplified-mod", "EM", file.list), em.dir, overwrite = TRUE)

## Start creating OM per iteration 
ss3sim::change_rec_devs(rdevs_array[-1,i], file.path(om.dir, "control.ss_new"), ctl_file_out = file.path(om.dir, "control.ss_new"))

dat <- r4ss::SS_readdat(
    file.path(om.dir, "data.ss"),
    verbose = FALSE
    )
ctl <- r4ss::SS_readctl(
    file.path(om.dir, "control.ss_new"),
    verbose = FALSE, use_datlist = TRUE, datlist = dat
    )

F_noncomm <- u_vec * F_vec_mat[i,]

ctl$F_Method <- 2
newctl <- ss3sim::change_f(years = list(1:75,1:75), fleets = c(1,2), fvals = list(F_vec_mat[i,], F_noncomm), ctl_list = ctl)
newctl$SR_parms$INIT[2] <- h_vec[2]
newctl$age_selex_types$Pattern <- 10
SS_writectl_3.30(newctl, outfile = file.path(om.dir, "control.ss"), overwrite = TRUE)

dat$catch$catch <- ifelse(dat$catch$catch == 0, 0, dat$catch$catch/dat$catch$catch)
##add dummy data for age comps
dat$agebin_vector <- seq(1,21)
dat$N_agebins <- length(dat$agebin_vector)
dat$N_ageerror_definitions <- 1
dat$ageerror <- matrix(data = c(rep(-1, 44), rep(0.001, 44)), nrow = 2, byrow = T)
dat$age_info <- data.frame("mintailcomp" = rep(-1,3), "addtocomp" = rep(1e-10,3), "combine_M_F" = 0, "CompressBins" = 0, "CompError" = 0, "ParmSelect" = 0, "minsamplesize" = 1)
dat$Lbin_method <- 3
dat$agecomp <- data.frame("Yr" = seq(1,75),
"Seas" = 1,
"FltSvy" = rep(3, 75),
"Gender" = 0,
"Part" = 0,
"Ageerr" = 1,
"Lbin_lo" = -1,
"Lbin_hi" = -1,
"Nsamp" = 10)
agecomp_mat <- as.data.frame(matrix(data = 1, nrow = nrow(dat$agecomp), ncol = dat$N_agebins))
dat$agecomp <- as.data.frame(cbind(dat$agecomp, agecomp_mat))
##add rows for earlier data for length comps 
xtra_lencomp <- data.frame("Yr" = seq(1,68),
"Seas" = 1,
"FltSvy" = rep(3, 68),
"Gender" = 0,
"Part" = 0,
"Nsamp" = 10)
lencomp_mat <-  as.data.frame(matrix(data = 1, nrow = nrow(xtra_lencomp), ncol = dat$N_lbins))
colnames(lencomp_mat) <- paste0("l", dat$lbin_vector)
xtra_lencomp <- cbind(xtra_lencomp, lencomp_mat)
dat$lencomp <- dat$lencomp %>% 
bind_rows(xtra_lencomp) %>% 
mutate(Seas = abs(Seas), FltSvy = abs(FltSvy)) %>% 
group_by(FltSvy) %>% 
arrange(Yr, FltSvy, .by_group = T) %>% 
as.data.frame()
SS_writedat_3.30(dat, outfile = file.path(om.dir, "data.ss"), overwrite = TRUE)

start <- SS_readstarter(file.path(om.dir, "starter.ss"))
start$datfile <- "data.ss"
start$ctlfile <- "control.ss"
start$run_display_detail <- 0
start$prior_like <- 0
start$last_estimation_phase <- 0
start$MCMCburn <- 0
start$MCMCthin <- 1
start$seed <- set.seed[1,1]
start$N_bootstraps <- 2
SS_writestarter(start, dir = om.dir, overwrite = TRUE)
#run ss to generate expected values for data
r4ss::run(dir = om.dir, exe = "ss_opt_win", extras = "-nohess", skipfinished = F)
## End of creating OM per iteration

## Start creating EMs per iteration
dat_exp <- SS_readdat_3.30(file.path(om.dir, "data_expval.ss"))
dat_samp <- ss3sim::sample_index(dat_exp, fleets = c(1,3), years = list(seq(1,75), seq(69,75)),
sds_obs = list(0.06, 0.25), seas = list(7,7))
dat_samp <- ss3sim::sample_catch(dat_samp) ##TODO: need to use u_vec to put in values for non_comm catch
dat_samp <- ss3sim::sample_lcomp(dat_samp, fleets = c(3), Nsamp = list(45), years = list(seq(69,75))) ##TODO: need to think more about Nsamp, cpar, and ESS
dat_samp <- ss3sim::sample_agecomp(dat_samp, fleets = c(3), Nsamp = list(45), years = list(seq(69,75))) 
SS_writedat_3.30(dat_samp, outfile = file.path(em.dir, "data.ss"), overwrite = TRUE)

ctl.em <- SS_readctl_3.30(file.path(em.dir, "control.ss_new"), use_datlist = TRUE, datlist = file.path(em.dir, "data.ss"))
#ctl.em$SR_parms #leave here as a reminder to change if needed if steepness is different
SS_writectl_3.30(ctl.em, outfile = file.path(em.dir, "control.ss"), overwrite = TRUE)

start <- SS_readstarter(file.path(em.dir, "starter.ss"))
start$datfile <- "data.ss"
start$ctlfile <- "control.ss"
start$N_bootstraps <- 1
SS_writestarter(start, dir = em.dir, overwrite = TRUE)

r4ss::run(dir = em.dir, exe = "ss_opt_win", skipfinished = F)

}

mods <- SSgetoutput(dirvec = c(
    dir.it.em[1],
    dir.it.om[1]
))
mods_sum <- SSsummarize(mods, verbose = F)
SSplotComparisons(mods_sum)
SSplotComparisons(mods_sum, legendlabels = c("em1", "em2", "em3", "em4", "em5", "om1", "om2", "om3","om4", "om5"))
length(which(mods_sum$maxgrad <= 1e-4))
SSplotData(mods$replist1)
SS_plots(mods$replist5)
##TODO: make plots to compare length and age comps across all iterations
mods_sum_sub <- SSsummarize(list(mods$replist2, mods$replist7))
SSplotComparisons(mods_sum_sub, legendlabels = c("EM", "OM"))
```



```{r}
dat_exp$CPUE %>% ggplot(aes(x = year, y = obs)) + geom_point() + facet_wrap(~index, scales = "free") + geom_line(data = dat_samp$CPUE, aes(x = year, y = obs), color = "blue") + geom_line(data = dat_samp2$CPUE, aes(x = year, y = obs), color ="orange")


mods_sum$SpawnBio %>% 
rename("em1" = "replist1",
"em2" = "replist2",
"em3" = "replist3",
"em4" = "replist4",
"em5" = "replist5", 
"om1" = "replist6",
"om2" = "replist7",
"om3" = "replist8", 
"om4" = "replist9",
"om5" = "replist10") %>%
pivot_longer(cols = -c("Label", "Yr")) %>%
separate(col = "name", into = c("Model", "iteration"), sep = "m") %>%
filter(Yr > 0) %>%
ggplot() +
geom_line(aes(x = Yr, y = value, group = interaction(iteration, Model), linetype = Model, color = Model), linewidth = 1.2) +
scale_linetype_manual(labels = c("EM", "OM"), values = c(1,2)) +
scale_color_manual(values = c(1,2), labels = c("EM", "OM")) +
facet_wrap(~iteration) +
theme_bw()


rep <- SS_output(file.path(main.dir, "Model_Runs", "fidi", "EM_1"))
SS_plots(rep)
rep <- SS_output(file.path(main.dir, "models", "03-simplified-mod"))
SS_plots(rep)


head(mods_sum$indices)
mods_sum$indices %>% 
select(Fleet, Fleet_name, Yr, Obs, Exp, name) %>% 
filter(str_detect(name, "replist1|replist6")) %>%
filter(str_detect(name, "replist10", negate = T)) %>%
ggplot(aes(x = Yr)) + 
geom_point(data = . %>% filter(str_detect(name, "replist6")), aes(y = Exp)) +  #OM observed data
geom_point(data = .%>% filter(str_detect(name, "replist1")), aes(y = Obs), color = "blue") +  #EM observed data
geom_line(data = .%>% filter(str_detect(name, "replist1")), aes(y = Exp), color = "blue") + 
facet_wrap(~Fleet_name, scales = "free") + theme_bw()

```

